---
title: "Hazard in Firebaugh"
author: "Samantha Liu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: false
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(sf)
library(leaflet)
library(tigris)
library(mapview)
library(jsonlite)
library(censusapi)
library(tidycensus)
library(StatMatch)
library(esri2sf)

Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")
census_api_key("c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

```
```{r block-groups, include = FALSE}
fresno_cbgs <- 
  block_groups("CA","Fresno County", cb = T, progress_bar = F) %>% 
  # filter(GEOID %in% paste0("0601900840", 11:16)) %>%   # Firebaugh block groups
  st_cast("MULTIPOLYGON") %>% 
  st_cast("POLYGON")

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = fresno_cbgs,
    color = "#e851a2",
    fillOpacity = 0.5,
    weight=1,
    label = ~GEOID
  )

```
```{r tracts, include = FALSE}
fresno_tracts <- 
  tracts("CA","Fresno County", cb = T, progress_bar = F) %>% 
  # filter(GEOID == "06019008401") %>%   # Firebaugh tract
  st_cast("MULTIPOLYGON") %>% 
  st_cast("POLYGON")

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = fresno_tracts,
    color = "#e851a2",
    fillOpacity = 0.5,
    weight=1,
    label = ~GEOID
  )
```

Building Data obtained from [Geofabrik](http://download.geofabrik.de/) makes many components of OSM available for easy download. After clicking a few levels down, you can find a [NorCal](http://download.geofabrik.de/north-america/us/california/norcal.html) page with a 700MB file for download with [these attributes](http://download.geofabrik.de/osm-data-in-gis-formats-free.pdf). Another option is Microsoft’s AI-generated [building footprints](https://github.com/Microsoft/USBuildingFootprints), which may be more or less accurate than OSM in different places. Their GitHub repo includes links to download the files as GeoJSONs by state, but California is massive, so a nice option is to use [county-level disaggregations](https://drive.google.com/drive/folders/1-XGvS25tQKKQ3HTqWjAfLJ4PaeXJ9yyY) done by [Patty Frontiera](https://dlab.berkeley.edu/blog/leave-only-footprints) at Berkeley. 
```{r building-parcel, include = FALSE}
fresno_parcels <- st_read("geographies/Fresno_Parcels_FeaturesToJSO.geojson")
fb_parcels <- 
  fresno_parcels %>% 
  filter(APN %in% paste0("007140", c("13ST", "04ST", "06T", "07ST"))) %>% 
  st_transform(4326)

# Microsoft AI-generating method
fresno_osm_bldg <- read_csv("data-raw/ca_06019_footprints.csv")
fb_osm_bldg <- 
  fresno_osm_bldg %>%
  st_as_sf(wkt = "WKT") %>%
  st_set_crs(4326) %>%
  rename(geometry = WKT) %>%
  .[fb_parcels, ] %>%
  rowid_to_column("ID") %>%
  filter(
    !ID %in% c(100, 78, 30, 23, 102, 27, 95, 89, 46, 94, 41, 67, 68, 61, 114, 7, 35, 86)
  )
```

```{r archive-building-parcel, include = FALSE, eval = FALSE}
# osm_bldg <- st_read("G:/Shared drives/SFBI/Data Library/OSM/gis_osm_buildings_a_free_1.shp")
# saveRDS(osm_bldg, "rds/osm_bldg.rds")

# Geofabrik method
osm_bldg <- readRDS("rds/osm_bldg.rds")
fb_osm_bldg <- 
  osm_bldg %>% 
  st_transform(st_crs(fresno_cbgs)) %>% 
  .[fresno_cbgs %>% filter(GEOID == "060190084011"),  ] 

temp <- tempfile()
download.file("https://www.co.fresno.ca.us/home/showpublisheddocument?id=12993", destfile = temp, mode = "wb")
fresno_parcels <- st_read(unzip(temp, "Fresno_Parcels.shp"))
unlink(temp)
```

## Hazard Data
Data Sources: [Cal Adapt](https://cal-adapt.org/)
```{r test-case}
projection <- "+proj=utm +zone=10 +ellps=GRS80 +datum=NAD83 +units=ft +no_defs"

test_cbg <- 
  fresno_cbgs %>% 
  filter(GEOID == "060190084011") %>%
  pull(geometry) %>% 
  st_transform(projection) %>% 
  st_simplify(
    dTolerance = ifelse(
      npts(fresno_cbgs[1, ]$geometry) > 100,
      400,
      0
    )
  ) %>% 
  st_transform(4326) %>% 
  st_as_text() %>% 
  str_replace_all(" ", "+")

test_exheat <- 
  fromJSON(
    paste0(
      "https://api.cal-adapt.org/api/series/tasmax_day_CanESM2_rcp85/exheat/?g=",
      test_cbg,
      "&stat=mean"
    )
  ) %>% 
  .$counts %>% 
  unlist() %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(
    year = rowname,
    exheat = "."
  ) %>% 
  mutate(
    year = substr(year, 1, 10) %>% as.Date()
  )

test_exheat %>% 
  ggplot(
    aes(
      x = year,
      y = exheat
    )
  ) +
  geom_line() +
  labs(
    x = "Year",
    y = "Extreme heat days",
    title = "CBG: 060190084011"
  )
```

Row numbers that result in a lexicon error: 46, 62, 108, 109, 134, 220
```{r}
exheat_full <- NULL

for(row in 1:nrow(fresno_cbgs)) {
  
  print(row)
  
  if(row %in% c(46, 62, 108, 109, 134, 220)) { 
    next 
  }
  
  temp <- 
    fromJSON(
      paste0(
        "https://api.cal-adapt.org/api/series/tasmax_day_CanESM2_rcp85/exheat/?g=",
        fresno_cbgs[row, ]$geometry %>%
          st_transform(projection) %>% 
          st_simplify(
            dTolerance = ifelse(
              npts(fresno_cbgs[row, ]$geometry) > 100,
              400,
              0
            )
          ) %>% 
          st_transform(4326) %>%
          st_as_text() %>% 
          str_replace_all(" ", "+"),
        "&stat=mean"
      )
    ) %>% 
    .$counts %>% 
    unlist() %>% 
    as.data.frame() %>% 
    rownames_to_column() %>% 
    rename(
      year = rowname,
      exheat = "."
    ) %>% 
    mutate(
      year = substr(year, 1, 10) %>% as.Date(),
      cbg = fresno_cbgs[row, ]$GEOID,
      tract = str_sub(cbg, 1, -2)
    )

  exheat_full <-
    exheat_full %>% 
    rbind(temp)
  
  if(row%%10 == 0) {
    print(row)
    saveRDS(exheat_full, "rds/exheat_full.rds")
  }
  
}

saveRDS(exheat_full, "rds/exheat_full.rds")
```

```{r}
exheat_full <- readRDS("rds/exheat_full.rds")

exheat_full %>% 
  group_by(year) %>% 
  summarize(
    exheat_mean = mean(exheat, na.rm = T),
    exheat_sd = sd(exheat, na.rm = T)
  ) %>%
  ggplot(
    aes(
      x = year
    )
  ) +
  geom_ribbon(
    aes(
      ymin = exheat_mean - exheat_sd,
      ymax = exheat_mean + exheat_sd
    ),
    fill = "grey"
  ) +
  geom_line(
    aes(
      y = exheat_mean
    )
  ) +
  labs(
    x = "Year",
    y = "Extreme heat days",
    title = "Fresno County CBGs"
  )
```

```{r}
exheat_2020s <- 
  exheat_full %>% 
  filter(year >= "2020-01-01" & year <= "2029-12-31") %>% 
  group_by(cbg) %>% 
  summarize(exheat = mean(exheat, na.rm = T)) %>% 
  left_join(fresno_cbgs %>% dplyr::select(cbg = GEOID)) %>% 
  st_as_sf()

heat_pal <- colorNumeric(
  palette = "Reds",
  domain = exheat_2020s$exheat
)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = exheat_2020s,
    fillColor = ~heat_pal(exheat),
    fillOpacity = 0.5,
    color = "white",
    weight = 0.5,
    label = ~paste0(cbg, ": ", exheat, " extreme heat days per year over the next decade"),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = exheat_2020s,
    pal = heat_pal,
    values = ~exheat,
    title = "Extreme heat days<br>per year over<br>next decade"
  )
```

## Exposure Data
Equity Analys on Heat Exposure - Assumes extreme heat affects everyone in a given census tract the same way

```{r}
bg_income <-
  getCensus(
    name = "acs/acs5",
    vintage = 2019,
    region = "block group:*",
    regionin = "state:06+county:019",
    vars = "group(B19001)"
  ) %>% 
  mutate(cbg = paste0(state, county, tract, block_group)) %>% 
  select(
    !c(GEO_ID, state, county, tract, block_group, NAME) & 
      !ends_with(c("EA", "MA", "M"))
  ) %>%
  pivot_longer(
    ends_with("E"),
    names_to = "variable",
    values_to = "estimate"
  ) %>%
  left_join(
    acs5_vars %>% 
      select(name, label), 
    by = c("variable" = "name")
  ) %>% 
  select(-variable) %>% 
  separate(
    label,
    into = c(NA,NA,"income"),
    sep = "!!"
  ) %>% 
  filter(!is.na(income)) %>% 
  mutate(
    income = case_when(
      income %in% c("Less than $10,000","$10,000 to $14,999","$15,000 to $19,999","$20,000 to $24,999") ~ "Less than $25,000",
      income %in% c("$25,000 to $29,999","$30,000 to $34,999","$35,000 to $39,999","$40,000 to $44,999","$45,000 to $49,999") ~ "$25,000 to $49,999",
      income %in% c("$50,000 to $59,999","$60,000 to $74,999") ~ "$50,000 to $74,999",
      TRUE ~ income
    )
  ) %>% 
  group_by(cbg, income) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE))

```

Extreme Heat Day buckets 250-300 should be lumped with 225-250 because of how small the total number of households they affect
```{r}
heat_exp_income <-
  bg_income %>% 
  left_join(
    exheat_full %>% 
      filter(year >= "2020-01-01" & year <= "2029-12-31") %>% 
      group_by(cbg) %>%
      summarize(exheat = sum(exheat)),
    by = c("cbg")
  ) %>% 
  filter(!is.na(exheat)) %>% 
  mutate(
    exheat_tier =
      case_when(
        exheat < 175 ~ "150-175",
        exheat < 200 ~ "175-200",
        exheat < 225 ~ "200-225",
        # exheat < 250 ~ "225-250",
        # exheat < 275 ~ "250-275",
        TRUE ~ "225-300"
      ) 
  ) %>% 
  group_by(income, exheat_tier) %>% 
  summarize(estimate = sum(estimate, na.rm = T)) %>% 
  mutate(
    income = income %>% factor(levels = rev(c("Less than $25,000", "$25,000 to $49,999", "$50,000 to $74,999", "$75,000 to $99,999", "$100,000 to $124,999", "$125,000 to $149,999", "$150,000 to $199,999", "$200,000 or more")))
  )
```

```{r}
heat_exp_income_stacked <-
  heat_exp_income %>% 
  rbind(
    heat_exp_income %>%
      group_by(income) %>%
      summarize(estimate = sum(estimate)) %>%
      mutate(exheat_tier = "Total")
  ) %>%
  ggplot() +
  geom_bar(
    aes(
      x = exheat_tier %>% factor(levels = c("150-175", "175-200", "200-225", "225-300", "Total")),
      y = estimate,
      fill = income
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Days of Extreme Heat",
    y = "Number of households",
    title = "Extreme Heat Day Exposure by Income",
    subtitle = "Fresno County - cumulated over 10 years",
    fill = "Household Income"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r}
heat_exp_income_fill <-
  heat_exp_income %>% 
  rbind(
    heat_exp_income %>%
      group_by(income) %>%
      summarize(estimate = sum(estimate)) %>%
      mutate(exheat_tier = "Total")
  ) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = exheat_tier %>% factor(levels = c("150-175", "175-200", "200-225", "225-300", "Total")),
      y = estimate,
      fill = income
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Days of Extreme Heat",
    y = "Number of households",
    title = "Extreme Heat Day Exposure by Income",
    subtitle = "Fresno County - cumulated over 10 years",
    fill = "Household Income"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

## Difference-in-Difference
```{r, eval = FALSE, inclue = FALSE}
bay_pumas <- readRDS("geographies/bay_pumas.rds")

pums_transit <- 
  2015:2017 %>% 
  map_dfr(function(year){
    get_pums(
      variables = c(
        "PUMA",
        "JWTR"
      ),
      state = "CA",
      year = year,
      survey = "acs1",
      recode = F
    ) %>% 
      mutate(
        bart = ifelse(
          JWTR %in% c("4","04"),
          PWGTP,
          0
        ),
        PUMA = PUMA %>% str_pad(5,"left","0")
      ) %>% 
      group_by(PUMA) %>% 
      summarize(
        pop = sum(PWGTP),
        bart = sum(bart),
        year = year
      )
  })

pums_transit_clean <-
  pums_transit %>% 
  select(-pop) %>% 
  pivot_wider(
    names_from = year,
    values_from = bart
  ) %>% 
  filter(PUMA %in% bay_pumas$PUMACE10)
```

```{r, eval = FALSE, inclue = FALSE}
obs_matrix <-
  pums_transit_clean %>% 
  select(`2015`,`2016`) %>% 
  as.matrix()

dist_matrix <- mahalanobis.dist(obs_matrix)

rownames(dist_matrix) <- pums_transit_clean$PUMA
colnames(dist_matrix) <- pums_transit_clean$PUMA

match <- 
  dist_matrix["08504", ] %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(
    PUMA = rowname,
    match = "."
  ) %>% 
  right_join(
    pums_transit_clean
  ) %>% 
  arrange(match) %>% 
  .[1:11, ] %>% 
  left_join(bay_pumas %>% select(PUMA = PUMACE10)) %>% 
  st_as_sf()

leaflet() %>% 
  addTiles() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = match[1, ],
    color = "red",
    label = ~PUMA
  ) %>% 
  addPolygons(
    data = match[-1, ],
    label = ~PUMA
  )
```

```{r, eval = FALSE, inclue = FALSE}
match_pumas <-
  match[-1,] %>% 
  st_set_geometry(NULL) %>% 
  select(-match) %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  ) %>%
  group_by(
    year
  ) %>% 
  summarize(
    bart = mean(bart),
    PUMA = "Similar PUMAs"
  )

treatment_puma <-
  match[1,] %>% 
  select(-match) %>% 
  st_set_geometry(NULL) %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  )

rbind(
  treatment_puma,
  match_pumas
) %>% 
  mutate(
    year = year %>% paste0(.,"-01-01") %>%  as.Date()
  ) %>% 
  ggplot(
    aes(
      x = year,
      y = bart,
      color = PUMA
    )
  ) +
  geom_line() +
  geom_vline(xintercept = 2017, linetype = "dashed") +
  labs(
    title = "Milpitas vs. control neighborhoods, BART ridership",
    x = "Year",
    y = "BART commute riders"
  )
```

```{r, eval = FALSE, inclue = FALSE}
transit_did <-
  match %>% 
  st_set_geometry(NULL) %>% 
  select(-match) %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  ) %>% 
  mutate(
    year = year %>% as.numeric(),
    time = ifelse(year == 2017, 1, 0),
    treated = ifelse(PUMA == "08504", 1, 0)
  )

did_reg <- lm(bart ~ treated*time, data = transit_did)

summary(did_reg)
```

## ACS Demographics
```{r ACS-functions}
acs5_vars <- readRDS("rds/acs5_vars.rds")

get_census_bg <- function(group) {
  getCensus(
    name = "acs/acs5",
    vintage = 2019,
    region = "block group:*",
    regionin = "state:06+county:019",
    vars = paste0("group(", group, ")")
  ) %>% 
    mutate(block_group = paste0(state, county, tract, block_group)) %>% 
    select(
      !c(GEO_ID, state, county, tract, NAME) & 
        !ends_with(c("EA", "MA", "M"))
    ) %>%
    pivot_longer(
      ends_with("E"),
      names_to = "variable",
      values_to = "estimate"
    ) %>%
    left_join(
      acs5_vars %>% 
        select(name, label), 
      by = c("variable" = "name")
    ) %>% 
    select(-variable) %>% 
    filter(block_group == "060190084011")
}

get_census_tract <- function(group) {
  getCensus(
    name = "acs/acs5",
    vintage = 2019,
    region = "tract:*",
    regionin = "state:06+county:019",
    vars = paste0("group(", group, ")")
  ) %>% 
  mutate(tract = paste0(state, county, tract)) %>%
  select(
    !c(GEO_ID, state, county, NAME) & 
      !ends_with(c("EA", "MA", "M"))
  ) %>%
  pivot_longer(
    ends_with("E"),
    names_to = "variable",
    values_to = "estimate"
  ) %>%
  left_join(
    acs5_vars %>% 
      select(name, label), 
    by = c("variable" = "name")
  ) %>% 
  select(-variable) %>% 
    filter(tract == "06019008401")
}

```

Total number of households recorded in Firebaugh according to ACS 2015-2019: 2321
Total number of occupied households recorded in Firebaugh according to ACS 2015-2019: 844
Total number recorded given by FHA: 2033
```{r pull-ACS-data}
# housing costs as % of income
fb_housing_costs <-
  get_census_tract("B25106") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure", "income", "% of income"),
    sep = "!!"
  ) %>% 
  filter(!is.na(`% of income`))

# HH Income - cbg level
fb_hhincome <-
  get_census_bg("B19001") %>% 
  separate(
    label,
    into = c(NA, NA, "income"),
    sep = "!!"
  ) %>% 
  filter(!is.na(income))

# aggregate vehicles available - cbg level
fb_vehicles <-
  get_census_bg("B25046") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure"),
    sep = "!!"
  ) %>% 
  filter(!is.na(`tenure`)) %>% 
  mutate(
    hh = c(342, 502), # obtained from fb_beds
    veh_per_hh = estimate / hh
  ) 

# bedrooms - cbg level
fb_beds <-
  get_census_bg("B25042") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure", "beds"),
    sep = "!!"
  ) %>% 
  filter(!is.na(beds))

# units per structure - cbg level
fb_units <-
  get_census_bg("B25032") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure", "units"),
    sep = "!!"
  ) %>% 
  filter(!is.na(units))

# hhsize by units in structure
fb_units_hhsize <-
  get_census_tract("B25124") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure", "HH Size", "units"),
    sep = "!!"
  ) %>% 
  filter(!is.na(units))

# total pop by units in structure - cbg level
fb_units_pop <-
  get_census_bg("B25033") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure", "units"),
    sep = "!!"
  ) %>% 
  filter(!is.na(units))

# heating fuel
fb_fuel <-
  get_census_tract("B25117") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure", "fuel"),
    sep = "!!"
  ) %>% 
  filter(!is.na(fuel))

# computer and internet access by educational attainment - cbg level
fb_internet_education <- 
  get_census_bg("B28006") %>% 
  separate(
    label,
    into = c(NA, NA, "education", "computer", "internet"),
    sep = "!!"
  ) %>% 
  mutate(
    internet = ifelse(computer == "No computer", "No computer", internet)
  ) %>% 
  filter(!is.na(internet))

# internet access by income - cbg level
fb_internet_income <- 
  get_census_bg("B28004") %>% 
  separate(
    label,
    into = c(NA, NA, "income", "internet"),
    sep = "!!"
  ) %>% 
  filter(!is.na(internet))

```

```{r archived-data, eval = FALSE, include = FALSE}
# number of rooms aggregated by owner or renter - cbg level
# redundant with fb_beds
fb_housing <-
  get_census_bg("B25022") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure"),
    sep = "!!"
  ) %>% 
  filter(!is.na(`tenure`))

# computer and internet access - cbg level
# redundant with fb_internet_education
fb_internet <- 
  get_census_bg("B28003") %>% 
  separate(
    label,
    into = c(NA, NA, "computer", "internet"),
    sep = "!!"
  ) %>% 
  mutate(
    internet = ifelse(computer == "No computer", "No computer", internet)
  ) %>% 
  filter(!is.na(internet))

# HH Income by owner or renter
# redundant by fb_housing_costs
fb_hhincome_tenure <-
  get_census_tract("B25118") %>% 
  separate(
    label,
    into = c(NA, NA, "tenure", "income"),
    sep = "!!"
  ) %>% 
  filter(!is.na(income))
```

```{r hhincome-tenure}
fb_housing_costs %>% 
  group_by(tract, tenure, income) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  mutate(
    income =
      income %>%
      substr(1, nchar(.) - 1) %>%
      factor(
        levels =
          c("Less than $20,000", "$20,000 to $34,999", "$35,000 to $49,999", "$50,000 to $74,999", "$75,000 or more")
      )
  ) %>%
  ggplot() +
  geom_bar(
    aes(
      x = income,
      y = estimate,
      fill = tenure
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household Income",
    y = "Proportion of Households",
    title = "Tenure by Household Income",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Tenure Classification"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r hhincome-percincome}
fb_housing_costs %>% 
  group_by(tract, `% of income`, income) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  mutate(
    income =
      income %>%
      substr(1, nchar(.) - 1) %>% 
      factor(
        levels =
          c("Less than $20,000", "$20,000 to $34,999", "$35,000 to $49,999", "$50,000 to $74,999", "$75,000 or more")
      ),
    `% of income` = `% of income` %>% factor(levels = c("Less than 20 percent", "20 to 29 percent", "30 percent or more"))
  ) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = income,
      y = estimate,
      fill = `% of income`
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household Income",
    y = "Proportion of Households",
    title = "Housing Costs as Percent of Income",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Percent of Income"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r hhsize-tenure}
fb_units_hhsize %>% 
  group_by(tract, tenure, `HH Size`) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = `HH Size`,
      y = estimate,
      fill = tenure
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household Size",
    y = "Proportion of Households",
    title = "Tenure by Household Size",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Tenure Classification"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r hhsize-units}
fb_units_hhsize %>% 
  group_by(tract, units, `HH Size`) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  mutate(units = ifelse(units == "1, detached  or attached:", "1, detached  or attached", units)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = `HH Size`,
      y = estimate,
      fill = units
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household Size",
    y = "Proportion of Households",
    title = "Units in Structure by Household Size",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Total Units in Structure"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r fuel-tenure}
fb_fuel %>% 
  group_by(tract, tenure, fuel) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = fuel,
      y = estimate,
      fill = tenure
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Fuel Source",
    y = "Proportion of Households",
    title = "Tenure by Fuel Source",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Tenure Classification"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

Median Income Bracket for the neighborhood is \$30,000 - \$34,999
Approximate number of vehicles per household for owned-homes: 2.62
Approximate number of vehicles per household for rented-homes: 0.92
```{r bedrooms-tenure}
fb_beds %>% 
  group_by(block_group, tenure, beds) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = beds,
      y = estimate,
      fill = tenure
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Number of Bedrooms",
    y = "Proportion of Households",
    title = "Tenure by Number of Bedrooms",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Tenure Classification"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r units-tenure-hh}
fb_units %>% 
  group_by(block_group, tenure, units) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = units,
      y = estimate,
      fill = tenure
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Total Units in Structure",
    y = "Total Households",
    title = "Tenure by Total Units in Structure",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Tenure Classification"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r units-tenure-population}
fb_units_pop %>% 
  group_by(block_group, tenure, units) %>% 
  summarize(estimate = sum(estimate, na.rm = TRUE)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = units,
      y = estimate,
      fill = tenure
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Total Units in Structure",
    y = "Total Number of People",
    title = "Tenure by Total Units in Structure",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Tenure Classification"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r education-internet}
fb_internet_education %>% 
  mutate(
    internet = ifelse(internet == "No computer", "", internet),
    comp_int = paste(computer, internet)
  ) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = education,
      y = estimate,
      fill = comp_int
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Educational Attainment",
    y = "Total Households",
    title = "Computer and Internet Access by Educational Attainment",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Computer and Internet Access"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

```{r hhincome-internet}
fb_internet_income %>% 
  mutate(
    income =
      income %>%
      substr(1, nchar(.) - 1) %>%
      factor(
        levels =
          c("Less than $10,000", "$10,000 to $19,999","$20,000 to $34,999", "$35,000 to $49,999", "$50,000 to $74,999", "$75,000 or more")
      )
  ) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = income,
      y = estimate,
      fill = internet
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Household Income",
    y = "Total Households",
    title = "Internet Access by Household Income",
    subtitle = "ACS 2015-2019 5-YR Estimates",
    fill = "Internet Access"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )
```

## Land-Use
```{r map-building-parcel}
leaflet() %>% 
  # addTiles(group = "OSM Base Map") %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = fb_parcels,
    color = "green",
    fillOpacity = 0.5,
    weight = 1,
    label = ~APN
  ) %>% 
  addPolygons(
    data = fb_osm_bldg,
    color = "#e851a2",
    fillOpacity = 0.5,
    weight = 1,
    label = ~ID
  )
```

```{r}
projection <- "+proj=utm +zone=10 +ellps=GRS80 +datum=NAD83 +units=ft +no_defs"

bldg_osm <- 
  fb_osm_bldg %>% 
  st_transform(projection)

parcels <-
  fb_parcels %>% 
  st_transform(projection)

yards <-
  parcels %>% 
  mutate(lot_area = st_area(.)) %>% 
  arrange(desc(lot_area)) %>% 
  st_difference(st_union(bldg_osm))

leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    data = yards %>% st_transform(4326),
    fillColor = "blue",
    color = "white",
    weight = 1,
    fillOpacity = 0.5
  )
```

```{r}
blocks <-
  parcels %>% 
  st_union()

street_edges <-
  blocks %>% 
  st_difference(blocks %>% st_buffer(-1)) %>% 
  st_intersection(parcels, .) %>% 
  filter(APN %in% yards$APN)

leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    data = yards %>% st_transform(4326),
    fillColor = "yellow",
    color = "white",
    weight = 1,
    fillOpacity = 0.5
  ) %>% 
  addPolygons(
    data = street_edges %>% st_transform(4326),
    fillColor = "brown",
    color = "brown",
    weight = 2
  )
```

```{r}
bldg_join <-
  bldg_osm %>% 
  st_centroid() %>% 
  st_join(parcels) %>% 
  st_set_geometry(NULL) %>% 
  left_join(bldg_osm %>% select(ID)) %>% 
  st_as_sf()

street_edges_buffered <-
  1:nrow(street_edges) %>% 
  map_dfr(function(x){
    
    temp_apn <- 
      street_edges[x,] %>% 
      pull(APN)
    
    temp_bldg <- 
      bldg_join %>% 
      filter(APN == temp_apn) %>% 
      st_union()
    
    nearest_distance <- 
      st_nearest_points(street_edges[x, ], temp_bldg) %>% 
      st_length()
    
    street_edges[x, ] %>% 
      st_buffer(nearest_distance)
    
  })

leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    data = yards %>% 
      st_transform(4326),
    fillColor = "yellow",
    color = "white",
    weight = 1,
    fillOpacity = 0.5
  ) %>% 
  addPolygons(
    data = street_edges_buffered %>% 
      st_transform(4326),
    fillColor = "brown",
    color = "brown",
    weight = 2
  )
```

```{r}
yards_setbacks <-
  parcels %>% 
  mutate(lot_area = st_area(.)) %>% 
  arrange(desc(lot_area)) %>%  
  st_buffer(-4) %>% 
  st_difference(st_union(bldg_osm %>% st_buffer(6))) %>% 
  st_difference(st_union(street_edges_buffered))

leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    data = yards_setbacks %>% st_transform(4326),
    fillColor = "blue",
    color = "white",
    weight = 1,
    fillOpacity = 0.5
  )
```

```{r}
buildable_area <-
  yards_setbacks %>% 
  st_buffer(-3.9) %>% 
  st_buffer(3.9, joinStyle = "BEVEL") %>%
  filter(!st_is_empty(.)) %>% 
  st_cast("MULTIPOLYGON") %>% 
  st_cast("POLYGON") %>% 
  mutate(
    buildable_area = st_area(.) %>% as.numeric()
  ) %>% 
  filter(buildable_area >= 160)

leaflet() %>% 
  addTiles() %>% 
  addPolygons(
    data = yards_setbacks %>% 
      st_transform(4326),
    fillColor = "red",
    color = "white",
    weight = 1,
    fillOpacity = 0.5
  ) %>% 
  addPolygons(
    data = buildable_area %>% 
      st_transform(4326),
    fillColor = "green",
    color = "white",
    weight = 1,
    fillOpacity = 1
  )
```

```{r}
buildable_area %>% st_union() %>% st_area()
```

